---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data




```{r loaddata, cache=TRUE}
data <- read.csv(unz("activity.zip", "activity.csv"))

#remove NA values from steps 
dataWOna <- subset(data, !is.na(data$steps))
```


## What is mean total number of steps taken per day?

####Histogram of the total number of steps taken each day
```{r totalstepsnumber , fig.width=10, fig.height=7}
library(ggplot2)

m <- ggplot(dataWOna, aes(x=date))
m <- m + geom_histogram(aes(weight = steps))
m <- m + theme_classic()
m <- m + theme(axis.text.x = element_text(angle = 90, hjust = 1))
m <- m + labs(
       x = "Dates",
       y = "Steps Number",
       title = "Total number of steps taken each day")
m
```

####Mean and median total number of steps taken per day
```{r mean_median}
library(plyr)
mean_median_steps <- ddply(dataWOna, .(date), summarise,
               mean = mean(steps),
               median = median(steps)
               )

print(mean_median_steps, row.names = FALSE)

```


## What is the average daily activity pattern?
```{r time_series_mean , fig.width=10, fig.height=7}
mean_interval_steps <- ddply(dataWOna, .(interval), summarise,
               mean_steps = mean(steps)
               )
max_mean_interval <- with(mean_interval_steps, interval[mean_steps == max(mean_steps)])

l <- ggplot(mean_interval_steps, 
       aes(interval, mean_steps))
l <- l + geom_line() 
l <- l + theme_classic()
l <- l + xlab("Day interval") 
l <- l + ylab("Mean steps")
l
```

`r max_mean_interval` interval  on average across all the days in the dataset, contains the maximum number of steps


## Imputing missing values
``` {r missing_values , fig.width=10, fig.height=7}
na_rows <- nrow(data[is.na(data$steps),])


df_mean <- merge(x = data, y = mean_interval_steps, by = "interval", all = TRUE)

#set steps to interval mean 
df_mean[is.na(df_mean$steps),]$steps <- df_mean[is.na(df_mean$steps),]$mean_steps 
df_mean$mean_steps <- NULL


#and combine dataset with mean and w/o mean
dataWOna$df_type <- "missing ignored"
df_mean$df_type <- "missing filled"

combinedData <- rbind(dataWOna, df_mean)

c <- ggplot(combinedData, aes(x=date, fill = df_type))
c <- c + geom_histogram(aes(weight = steps), position="dodge")
c <- c + theme_classic()
c <- c + theme(axis.text.x = element_text(angle = 90, hjust = 1))
c <- c + labs(
       x = "Dates",
       y = "Steps Number",
       title = "Total number of steps taken each day")
c <- c + theme(legend.position="bottom")
c
```



#### The is no impact of imputing missing data except the dates that all the data was missing


## Are there differences in activity patterns between weekdays and weekends?

```{r mean_median_steps_combined}
mean_median_steps_combined <- merge (mean_median_steps,
                ddply(df_mean, .(date), summarise, mean_wo_missing = mean(steps), median_wo_missing = median(steps)), 
                by="date", all = TRUE)

print(mean_median_steps_combined, row.names = FALSE)
```

```{r weekend, fig.width=10, fig.height=7}
library(timeDate)

df_mean$isWeekDay <- factor(isWeekday(as.POSIXct(df_mean$date, format="%Y-%m-%d"), wday = 1:5), levels=c(TRUE, FALSE), labels=c("weekday", "weekend"))


df_mean_interval_steps <- ddply(df_mean, .(interval, isWeekDay), summarise,
               mean_steps = mean(steps)
               )

l <- ggplot(df_mean_interval_steps, 
       aes(interval, mean_steps))
l <- l + geom_line() 
l <- l + theme_classic()
l <- l + xlab("Day interval") 
l <- l + ylab("Mean steps")
l <- l + facet_wrap(~isWeekDay, nrow = 2)
l

```

